---
output:
  github_document:
    html_preview: false
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = FALSE
)

run_live_api <- nzchar(Sys.getenv("GRANTS_GOV_API_KEY")) &&
  identical(tolower(Sys.getenv("GRANTSGOV_README_LIVE")), "true")
```

# grantsgov

<!-- badges: start -->
[![R-CMD-check](https://github.com/muschellij2/grantsgov/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/muschellij2/grantsgov/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/muschellij2/grantsgov/graph/badge.svg)](https://app.codecov.io/gh/muschellij2/grantsgov)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![License: MIT](https://img.shields.io/badge/license-MIT-blue.svg)](LICENSE.md)
<!-- badges: end -->

`grantsgov` is an R client for the Simpler Grants.gov API at
<https://api.simpler.grants.gov/docs#/>. It uses `httr2` for requests and maps
the documented core endpoints:

- `POST /v1/opportunities/search`
- `GET /v1/opportunities/{opportunity_id}`
- `POST /v1/extracts`

Set your API key in the `GRANTS_GOV_API_KEY` environment variable:

```{r api-key-example, eval = FALSE}
Sys.setenv(GRANTS_GOV_API_KEY = "your-api-key")
```

## Installation

```{r install-example, eval = FALSE}
# install.packages("pak")
pak::pak("muschellij2/grantsgov")
```

## Load the package

```{r load-package, message = FALSE}
if (file.exists("DESCRIPTION") && dir.exists("R")) {
  pkgload::load_all(quiet = TRUE)
} else {
  library(grantsgov)
}
```

## Inspect endpoint metadata

These examples run without an API key because they only inspect local package
metadata.

```{r endpoint-metadata}
grant_base_url()
names(grant_endpoints()$endpoints)
grant_endpoints()$endpoints$health
grant_search_options()$sort_by
grant_extract_options()$extract_type
grant_rate_limit_headers()
```

## Build request components

The helper constructors create the nested request-body structures used by the
API.

```{r request-components}
filters <- list(
  opportunity_status = grant_filter_one_of(c("posted", "forecasted")),
  applicant_type = grant_filter_one_of("nonprofits"),
  close_date = grant_filter_date_range(
    start_date = Sys.Date(),
    end_date = Sys.Date() + 30
  ),
  award_ceiling = grant_filter_number_range(max = 1000000)
)

pagination <- grant_pagination(
  page_offset = 1,
  page_size = 5000,
  sort_order = grant_sort("close_date", "ascending")
)

str(filters)
pagination
```

## Search opportunities

This chunk runs when `GRANTS_GOV_API_KEY` is set and
`GRANTSGOV_README_LIVE=true`.

```{r search-opportunities, eval = run_live_api}
results <- grant_search_opportunities(
  query = "education",
  filters = filters,
  pagination = pagination
)

length(results$data)
```

To request CSV content instead of JSON:

```{r search-opportunities-csv, eval = run_live_api}
csv_text <- grant_search_opportunities(
  filters = list(opportunity_status = grant_filter_one_of("posted")),
  pagination = grant_pagination(page_size = 100),
  format = "csv"
)

substr(csv_text, 1, 200)
```

To collect more than one page, use `grant_paginate()` or the endpoint-specific
wrapper `grant_search_all_opportunities()`.

```{r paginate-search, eval = FALSE}
all_results <- grant_search_all_opportunities(
  query = "education",
  filters = filters,
  page_size = 5000,
  sort_order = grant_sort("close_date", "ascending")
)

length(all_results)
attr(all_results, "pagination_info")
```

## Recent opportunities

Use `grant_recent_opportunities()` to paginate through opportunities posted in
the last day or last week. The API exposes this as a `post_date` filter.

```{r recent-opportunities, eval = FALSE}
last_day <- grant_recent_opportunities("day")
last_week <- grant_recent_opportunities("week")

length(last_day)
length(last_week)
```

## Get opportunity details

This example needs a real opportunity UUID from the API.

```{r get-opportunity, eval = FALSE}
opportunity <- grant_get_opportunity(
  "12345678-1234-1234-1234-123456789012"
)
```

## List, download, and read extracts

The metadata request runs when `GRANTS_GOV_API_KEY` is set and
`GRANTSGOV_README_LIVE=true`. Use the CSV extract type when you want to read the
downloaded data into R.

```{r list-extracts, eval = run_live_api}
extracts <- grant_list_extracts(
  filters = list(
    extract_type = "opportunities_csv",
    created_at = grant_filter_date_range(Sys.Date() - 30, Sys.Date())
  ),
  pagination = grant_pagination(
    page_size = 5000,
    sort_order = grant_sort("created_at")
  )
)

length(extracts$data)
```

`grant_download_extract()` defaults to a temporary `.csv` file. Pass
`expected_file_size = extract$file_size_bytes` to verify the downloaded byte
count against the API metadata.

```{r download-extract, eval = FALSE}
extract_file <- grant_download_extract(
  extracts$data[[1]],
  expected_file_size = extracts$data[[1]]$file_size_bytes
)

extract_file
```

`grant_read_extract()` downloads a CSV extract, reads it with `readr::read_csv()`,
warns if readr reports parsing problems, and stores the downloaded file path in
`attr(data, "file")`.

```{r read-extract, eval = FALSE}
extract_data <- grant_read_extract(
  extracts$data[[1]],
  expected_file_size = extracts$data[[1]]$file_size_bytes
)

attr(extract_data, "file")
```

Use `grant_list_all_extracts()` to automatically request each page of extract
metadata.

```{r paginate-extracts, eval = FALSE}
all_extracts <- grant_list_all_extracts(
  filters = list(extract_type = "opportunities_csv"),
  page_size = 5000
)

length(all_extracts)
```

## Errors and rate limits

HTTP errors include the status code, API-provided error message when available,
and any returned rate-limit headers such as `Retry-After`,
`X-RateLimit-Remaining`, or `RateLimit-Reset`. A `429` response is reported with
retry/backoff guidance.
