---
title: "Using grantsgov"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using grantsgov}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = FALSE
)

run_live_api <- nzchar(Sys.getenv("GRANTS_GOV_API_KEY")) &&
  identical(tolower(Sys.getenv("GRANTSGOV_VIGNETTE_LIVE")), "true")
```

```{r setup, message = FALSE}
if (file.exists("../DESCRIPTION") && dir.exists("../R")) {
  pkgload::load_all(path = "..", quiet = TRUE)
} else {
  library(grantsgov)
}
```

The `grantsgov` package wraps the Simpler Grants.gov API. It helps you search
grant opportunities, retrieve details for a specific opportunity, and list bulk
extract metadata.

Most API requests require an API key. The package reads it from
`GRANTS_GOV_API_KEY`.

```{r api-key, eval = FALSE}
Sys.setenv(GRANTS_GOV_API_KEY = "your-api-key")
```

Live API examples in this vignette run only when both environment variables are
set:

```{r live-env, eval = FALSE}
Sys.setenv(
  GRANTS_GOV_API_KEY = "your-api-key",
  GRANTSGOV_VIGNETTE_LIVE = "true"
)
```

## Discover available endpoints and options

The package documents the API paths, auth header, supported search filters, sort
fields, extract filters, and rate-limit headers checked by the error handler.

```{r endpoint-options}
grant_endpoints()$endpoints
grant_endpoints()$endpoints$health
grant_search_options()$filters
grant_search_options()$sort_by
grant_extract_options()
grant_rate_limit_headers()
```

## Common use case: find open education opportunities

A common workflow is to search for open or forecasted opportunities, filter by
deadline and award size, then sort the results by close date.

```{r build-education-search}
education_filters <- list(
  opportunity_status = grant_filter_one_of(c("posted", "forecasted")),
  close_date = grant_filter_date_range(
    start_date = Sys.Date(),
    end_date = Sys.Date() + 90
  ),
  award_ceiling = grant_filter_number_range(min = 50000, max = 2000000)
)

education_pagination <- grant_pagination(
  page_offset = 1,
  page_size = 5000,
  sort_order = grant_sort("close_date", "ascending")
)

str(education_filters)
education_pagination
```

Run the search with `grant_search_opportunities()`.

```{r run-education-search, eval = run_live_api}
education_results <- grant_search_opportunities(
  query = "education",
  filters = education_filters,
  pagination = education_pagination
)

length(education_results$data)
```

## Common use case: find NIH grants in a subject area

To search for NIH opportunities in a particular area, use a subject keyword and
filter to the Department of Health and Human Services top-level agency. NIH
opportunities are returned under agency names such as "National Institutes of
Health".

```{r nih-area-search, eval = FALSE}
nih_cancer <- grant_search_opportunities(
  query = "cancer",
  filters = list(
    top_level_agency = grant_filter_one_of(list("HHS")),
    opportunity_status = grant_filter_one_of(c("posted", "forecasted"))
  ),
  pagination = grant_pagination(
    page_size = 25,
    sort_order = grant_sort("close_date", "ascending")
  )
)

length(nih_cancer$data)
```

Older Grants.gov-style parameters map to the current API like this:

```{r old-style-search, eval = FALSE}
keyword <- "education"
limit <- 25

results <- grant_search_opportunities(
  query = keyword,
  pagination = grant_pagination(
    page_offset = 1, # startRecordNum = 0
    page_size = limit, # rows
    sort_order = grant_sort("close_date", "ascending") # closeDate ASC
  )
)
```

## Retrieve details for one opportunity

Search results include opportunity identifiers. Pass one of those identifiers to
`grant_get_opportunity()` to retrieve the full record.

```{r get-first-opportunity, eval = run_live_api}
first_id <- education_results$data[[1]]$opportunity_id
first_opportunity <- grant_get_opportunity(first_id)

names(first_opportunity$data)
```

If you already know an opportunity ID, call the function directly.

```{r get-known-opportunity, eval = FALSE}
grant_get_opportunity("12345678-1234-1234-1234-123456789012")
```

## Paginate through all search results

After checking a first response, pass it to `grant_fetch_all()` to collect the
remaining pages with the same query, filters, sort order, and page size.

```{r fetch-all-search, eval = FALSE}
first_education_page <- grant_search_opportunities(
  query = "education",
  filters = education_filters,
  pagination = grant_pagination(
    page_size = 25,
    sort_order = grant_sort("close_date", "ascending")
  )
)

all_education_results <- grant_fetch_all(first_education_page)

length(all_education_results)
attr(all_education_results, "pagination_info")
```

If you request a different `page_size`, `grant_fetch_all()` refetches from page
1 because API page numbers depend on page size.

```{r fetch-all-search-large-page, eval = FALSE}
all_education_results <- grant_fetch_all(first_education_page, page_size = 5000)
```

You can also use `grant_paginate()` for any endpoint function with a
`pagination` argument, or use `grant_search_all_opportunities()` for
search-specific defaults.

```{r paginate-search, eval = FALSE}
all_education_results <- grant_search_all_opportunities(
  query = "education",
  filters = education_filters,
  page_size = 5000,
  sort_order = grant_sort("close_date", "ascending")
)

length(all_education_results)
attr(all_education_results, "pagination_info")
```

## Common use case: find recently posted opportunities

Use `grant_recent_opportunities()` to collect all opportunities posted in the
last day or last week. The API exposes this as a `post_date` filter.

```{r recent-opportunities, eval = FALSE}
last_day <- grant_recent_opportunities("day")
last_week <- grant_recent_opportunities("week")

length(last_day)
length(last_week)
```

## Export search results as CSV text

The search endpoint can return CSV content. This is useful when a downstream
workflow expects a delimited file.

```{r search-csv, eval = run_live_api}
csv_text <- grant_search_opportunities(
  query = "education",
  filters = education_filters,
  pagination = education_pagination,
  format = "csv"
)

cat(substr(csv_text, 1, 300))
```

## Common use case: find recent bulk extracts

Bulk extracts are useful when you want a complete snapshot rather than a page of
search results.

```{r build-extract-search}
extract_filters <- list(
  extract_type = "opportunities_csv",
  created_at = grant_filter_date_range(Sys.Date() - 30, Sys.Date())
)

extract_pagination <- grant_pagination(
  page_offset = 1,
  page_size = 5000,
  sort_order = grant_sort("created_at", "descending")
)

extract_filters
extract_pagination
```

List available extract metadata with `grant_list_extracts()`.

```{r list-extracts, eval = run_live_api}
extracts <- grant_list_extracts(
  filters = extract_filters,
  pagination = extract_pagination
)

length(extracts$data)
```

Download one extract by passing the metadata record or a download URL to
`grant_download_extract()`. The default output path is a temporary `.csv` file.
Pass `expected_file_size = extract$file_size_bytes` to verify the downloaded
byte count against the API metadata.

```{r download-extract, eval = FALSE}
extract_file <- grant_download_extract(
  extracts$data[[1]],
  expected_file_size = extracts$data[[1]]$file_size_bytes
)

extract_file
```

Read a CSV extract with `grant_read_extract()`. It downloads the file, reads it
with `readr::read_csv()`, warns if readr reports parsing problems, and stores
the downloaded file path in `attr(data, "file")`.

```{r read-extract, eval = FALSE}
extract_data <- grant_read_extract(
  extracts$data[[1]],
  expected_file_size = extracts$data[[1]]$file_size_bytes
)

attr(extract_data, "file")
```

Use `grant_list_all_extracts()` to automatically request every page of extract
metadata.

```{r paginate-extracts, eval = FALSE}
all_extracts <- grant_list_all_extracts(
  filters = list(extract_type = "opportunities_csv"),
  page_size = 5000
)

length(all_extracts)
```

## Error handling and rate limits

When the API returns an error, the package reports the HTTP status, a short
status-specific hint, the API's message body when available, and any rate-limit
headers returned by the gateway. A `429` response includes retry/backoff
guidance and rate-limit metadata such as `Retry-After`,
`X-RateLimit-Remaining`, or `RateLimit-Reset` when present.

```{r key-error, error = TRUE}
withr::with_envvar(
  c(GRANTS_GOV_API_KEY = NA),
  grant_api_key()
)
```
